---
title: "Getting Started with pllr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with pllr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Installation

Install from GitHub using devtools:

```{r install, eval = FALSE}
# install.packages("devtools")
devtools::install_github("WalrusQuant/pllr")
```

Then load the package:

```{r load, eval = FALSE}
library(pllr)
```

## Overview

`pllr` wraps the Premier Lacrosse League (PLL) stats API and returns results as tibbles with `snake_case` column names. The package provides six functions covering standings, player statistics, game events, the draft, and free agency.

All column names are automatically converted from the API's `camelCase` to `snake_case`.

## Functions

### `pll_standings()` — Team Standings

Returns team standings for a given season. Set `champ_series = TRUE` for Champ Series standings or `FALSE` for regular season.

```{r standings, eval = FALSE}
standings <- pll_standings(year = 2025, champ_series = TRUE)
# # A tibble: 14 × 19
# # team_id  full_name  location  wins  losses  ties  scores  scores_against ...
```

Key columns: `team_id`, `full_name`, `location`, `seed`, `wins`, `losses`, `ties`, `scores`, `scores_against`, `score_diff`, `conference`, `conference_seed`.

### `pll_events()` — Game Schedule & Results

Returns the game schedule and results for a season. Optionally include Champ Series and WLL events.

```{r events, eval = FALSE}
events <- pll_events(year = 2025, include_cs = TRUE, include_wll = TRUE)
# # A tibble: ~80 × 64
# # id  event_id  league  start_time  venue  broadcaster  home_full_name  away_full_name ...
```

Home team columns are prefixed with `home_` and away team columns with `away_` (e.g., `home_full_name`, `away_score`, `away_url_logo`).

### `pll_player_stats()` — Player Statistics

Returns per-player season statistics. Use `season_segment` to choose between `"champseries"` and `"regular"`.

```{r player_stats, eval = FALSE}
stats <- pll_player_stats(year = 2025, season_segment = "champseries")
# # A tibble: ~200 × 70
# # official_id  first_name  last_name  position  goals  assists  points ...
```

Player identity columns include `official_id`, `first_name`, `last_name`, `slug`, `jersey_num`, `position`, `experience`. Team columns are prefixed with `team_` (e.g., `team_full_name`, `team_location_code`). Stat columns include `goals`, `assists`, `points`, `shots`, `shot_pct`, `saves`, `save_pct`, `faceoffs_won`, `faceoff_pct`, `ground_balls`, `turnovers`, `caused_turnovers`, `gaa`, `plus_minus`, and many more.

### `pll_free_agents()` — Free Agent Data

Returns free agent data for a given year.

```{r free_agents, eval = FALSE}
free_agents <- pll_free_agents(year = 2026)
# # A tibble: ~50 × 14
# # name  position  age  prev_team_id  new_team_id  new_contract_status ...
```

Columns include `name`, `first_name`, `last_name`, `slug`, `age`, `experience`, `position`, `status`, `official_id`, `new_contract_status`, `new_team_id`, `prev_team_id`, `thirty_percent_threshold_met`.

Note: `pll_free_agents()` uses the GraphQL API endpoint; all other functions use the REST API.

### `pll_draft_order()` — Draft Pick Order

Returns the official draft pick order for a given year and league.

```{r draft_order, eval = FALSE}
draft <- pll_draft_order(year = 2025, league = "PLL")
# # A tibble: ~60 × 11
# # id  year  round  round_pick  overall_pick  team_id  pick_player_name ...
```

Pick-specific columns are prefixed with `pick_`: `pick_player_name`, `pick_college`, `pick_position`, `pick_player_id`. Pick columns are `NA` for unfilled slots.

### `pll_draft_predictions()` — Mock Draft Rankings

Returns analyst mock draft rankings (not official picks) for a given year.

```{r draft_predictions, eval = FALSE}
predictions <- pll_draft_predictions(year = 2026)
# # A tibble: ~30 × 13
# # id  year  analyst_name  player_name  position  college  overall_rank  analysis ...
```

Columns include `id`, `year`, `analyst_name`, `player_name`, `image_url`, `position`, `college`, `college_logo`, `overall_rank`, `position_rank`, `change`, `analysis`, `league`.

Use `pll_draft_order()` for official selections after the draft has occurred.

## Notes

**Column naming:** All returned tibble columns use `snake_case`. The underlying API uses camelCase; the package converts all names automatically.

**Data availability:** Season data (standings, player stats, events) is only available during or after the season has been played. Querying a future or in-progress season may return partial or empty results.

**Draft predictions vs. draft order:** `pll_draft_predictions()` returns analyst mock draft rankings. Use `pll_draft_order()` for official draft selections.
